Education and the Rate of Entrepreneurship - Data Gathering and Analysis

### 2014-11-14 GEMDataSource.Rmd
Formatting and sims

---
title: "The Country-level Impact of Educational Features on the Rate of Entrepreneurship - Data Gathering and Analysis"
author: "Samuel Dundon and Thierry Racaud"
date: "November 14, 2014"
output: html_document
---

@@

# Library Repmis, Tidy, Reshape2, and plyr packages
library(repmis)
library(tidyr)
library(reshape2)
library(plyr)
library(WDI)
library(stargazer)
library(Zelig)

@@

# Check data to confirm change
summary(CompleteDataset)

# Run a linear regression
R1 <- lm(PercentIntention ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

@@

# Calculate confidence intervals of R1
confint(R1)

# Run a linear regression of R2
R2 <- lm(PercentNascent ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

@@

# Calculate confidence intervals of R2
confint(R2)

# Run a linear regression of R3
R3 <- lm(PercentOwner ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

# Obtain summary statistics of R3
summary(R3)

# Calculate confidence intervals of R3
confint(R3)

# Plot graph
car::scatterplotMatrix(CompleteDataset)
```

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis'} 
# Create regression output table
labels <- c('Percent Perceived', 'Percent Fear',
            'Annual GDP Growth', 'Percent Opportunities', 'Percent Fear x Percent Opportunities','(Intercept)')

stargazer::stargazer(R1, R2, R3, covariate.labels = labels, 
                     title = 'Linear Regression Estimates', digits = 2, type = 'html')

@@

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis'}
library(Zelig)

Z1 <- zelig(PercentIntention ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, cite = FALSE,
            data = CompleteDataset, model = 'ls', param = function(z.out))

z5 <- zls$new()

summary(Z1)

Z1$sim()

sim(setz1)

### Research Question

To what extent do the behavioural characteristics of education systems - individualistic, group-oriented, self-esteem reinforcing - affect the rate of entrepreneurship in OECD countries?

### Estimation Strategy

Education system / Teaching reinforcing self-esteem ---> higher degree of individual self-esteem ---> higher probability an individual will become an entrepreneur ---> higher rate of entrepreneurship at a country-level.

#### Data Collection

To find out, we need to gather quantitative data about entrepreneurship on the one hand, and some measures of these behavioural characteristics on the other hand.

##### The rate of entrepreneurship at country-level is the variable we intend to explain

We intend to look into three dimensions of entrepreneurship. We intend to look at established entrepreneurship in the country, whether it is successful or not, nascent entrepreneurship (less than three months), again whether it is successful or not, and entrepreneurial intention. We will study them successively as we may well observe different effects.

To measure these three dimensions of entrepreneurship, we will use three [Global Entrepreneurship Monitor (GEM)](http://www.gemconsortium.org) indicators:

##### Degree of Self-esteem reinforcement within the education system will be our key explanatory variable

Inglehart showed how personal values are shaped before the age of 20, we will try to show how compulsory education (from 6 to 16 years of age) does shape the individual, especially in terms of self-esteem. Of course, is is not easy to quantify how supportive of students' self-esteem a given education system can be and we are still looking for the most adequate measure.

The closest data we found so far is one of the six items / questions that GEM uses to measure the importance of entrepreneurial education at country-level in its NES Survey. Item D01 states: "In my country, teaching in primary and secondary education encourages creativity, self-sufficiency, and personal initiative".

For now, we use two GEM indicators that relate to self-esteem and confidence.

We look at the average "perceived capabilities" of an individual (18-64) of a given country to start a business. The data consists of average percentage at country-level.

We also take into account how the "fear of failure" could deter the average individual (18-64) of a country to become an entrepreneur. The data consists of average percentage at country-level.

##### We will include measures to check how individualistic and group-oriented a given education system is

Literature generally admits that self-confidence partly depends on the integration in social networks. Therefore, as we try to determine how supportive of students' self-esteem a given education system, we will look at how individualistic or group-oriented it is. To quantify such a behavioural feature is not easy, which is why we will have to rely on another measure: how individualistic or group-oriented an average student is by the end of compulsory education. Since OECD PISA collects data about students at the end of compulsory education, we will rely on "the sense of belonging" it defines.

##### We will control for classic determinants of entrepreneurship

We will intend to control for the classic determinants of entrepreneurship, that is for economic and structural determinants. More specifically, we will control for nine key entrepreneurship conditions as defined in the GEM [National Experts Survey (NES)](http://http://www.gemconsortium.org/Data-Collection) that may affect the rate of entrepreneurship. These nine conditions are the following:

We are not yet fully sure on how best to control for all these variables, especially as GEM NES data does not seem to be available in dataset format as such. We may need to build up our own dataset, collecting data from the many national GEM NES reports. We will work on this in the coming weeks.

As for now, as a substitute to all these conditions together, we use GEM indicator called "Perceived Opportunities". For each country, it is the average percentage of individuals who believe there are opportunities to start a business in the area they live in. This is not equivalent to controlling for all the conditions above but it actually tells about all of them from the potential entrepreneur's point of view. Since we look at average individual self-esteem in every OECD country, we believe it is consistent to look at the opportunities perceived by the average 18-64-year-old individual of these countries.

Besides these nine conditions and the GEM "perceived opportunities" measure, we also intend to control for the economic outlook at the time of data collection which is why we chose to include GDP annual growth rate for each country from 2001 until 2013. This is our only data that does not originates from the GEM collection.

#### Data Scraping and Cleaning

In order to scrape our data, we used repmis, tidyr, reshape2, plyr, WDI, stargazer and Zelig packages. We then imported one dataset per variable mentioned above:

GEM Established Entrepreneurship Rate

GEM Nascent Entrepreneurship Rate

GEM Entrepreneurial Intention

GEM Perceived Capabilities

GEM Fear of Failure

GEM Perceived Opportunities

GDP Annual Growth Rate

.

>>>>

---
title: Education and the Rate of Entrepreneurship - Data Gathering and Analysis
author: "Samuel Dundon and Thierry Racaud"
date: "November 14, 2014"
output: html_document
---
***

#### Research Question

To what extent do the behavioral characteristics of education systems - individualistic, group-oriented, self-esteem reinforcing - affect the rate of entrepreneurship in OECD countries?  
<br>

Education system / Teaching reinforcing self-esteem ---> higher degree of individual self-esteem ---> higher probability an individual will become an entrepreneur ---> higher rate of entrepreneurship at a country-level.
<br>

##### Dependent Variables: Explaining the Country-level Rate of Entrepreneurship

We have examined three dimensions of entrepreneurship to prepare for this stage of the project. We have analyzed established entrepreneurship in the country, nascent entrepreneurship (less than three months), and entrepreneurial intention. We will study them successively as we may well observe different effects.

To measure these three dimensions of entrepreneurship, we have used three [Global Entrepreneurship Monitor (GEM)](http://www.gemconsortium.org) indicators:

##### Key Explanatory Variable: Degree of Self-esteem Reinforcement within the Education System

As Inglehart has shown how personal values are shaped before the age of 20, we will try to show how compulsory education (from 6 to 16 years of age) shapes individuals, especially in terms of self-esteem. Of course, is is not easy to quantify how supportive of students' self-esteem a given education system can be and we are still looking for the most adequate measure.

To begin with, we decided to incorporate to GEM indicators that relate to self-esteem and confidence. First, included GEM survey data on the average "perceived capabilities" of an individual (18-64) of a given country to start a business. The data consists of average percentage at country-level. Second, we included additional GEM data on how the "fear of failure" could deter the average individual (18-64) of a country to become an entrepreneur. The data consists of average percentage at country-level.

Going forward, we are considering adding one of the six questions used by GEM to measure the importance of entrepreneurial education at country-level in its NES Survey. Item D01 states: "In my country, teaching in primary and secondary education encourages creativity, self-sufficiency, and personal initiative."

##### Control Variables: Controlling for Classic Determinants of Entrepreneurship

To control for the general state of the economy globally and in each country, we included Annual GDP Growth from the World Bank as a control variable in our regression equations.  We first considered including real GDP, but decided to include growth to more accurately capture fluctuation in the economy.  In addition, we included a GEM indicator called "Perceived Opportunities." For each country, it is the average percentage of individuals who believe there are opportunities to start a business in the area they live in. Since we look at average individual self-esteem in every OECD country, we believe it is appropriate to look at the opportunities perceived by the average 18-64-year-old individual of these countries.

In the future, we may include lagged GDP growth in recognition of the time taken for governments and individuals to react to the state of the economy, and are planning to include several control variables from GEM. Specifically, we intend to control for the classic determinants of entrepreneurship -- economic and structural determinants. To do so, we will add the following nine key entrepreneurship conditions as defined in the GEM [National Experts Survey (NES)](http://http://www.gemconsortium.org/Data-Collection) that may affect the rate of entrepreneurship.

The following equations were used to determine the relationship between our dependent and explanatory variables:

<br>

1. $PercentIntention = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

2. $PercentNascent = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

3. $PercentOwner = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$  
<br>

The variables *Percent Fear* and *Percent Opportunities* were included as an interaction term $PercentFear*PercentOpportunities$ because of the design of the survey. PercentFear is:

>The Percentage of 18-64 population with positive perceived opportunities who indicate that fear of failure would prevent them from setting up a business.

<br>

#### Data Collection and Cleaning

To collect the data necessary to run these regression equations, we downloaded *.csv* data from GEM and used the WDI API to obtain the necessary Annual GDP growth of OECD Countries.

**Necessary R Packages**

The following packages were needed to complete the data collection portion of this assignment:

* repmis  
* tidyr  
* reshape  
* plyr  
* WDI

**GEM Data**

GEM data was embedded in an interactive database and was difficult to extract automatically without writing a program. To simplify this process, we downloaded the necessary *.csv* files and saved them in our GitHub repository (to reduce complexity, we deleted the first three unnecessary rows before saving).  We then linked to the RAW version of the files and added the code required to load them from *R*.  The repmis package is required to load these documents from the repository.

Cleaning the GEM data required deleting unnecessary characters in the year columns, changing the years from columns to rows (changing the data from wide to long), and changing the column names to describe which data was being analyzed.  In the process of doing this we created new data sets for each of these survey questions, which were called C*dataset* (clean followed by the data set name). This was done for each of the six GEM data sets before merging these sets two at a time to create the data set *FinalGEM*.

For all 33 OECD countries we should have 13 observations, one per year from 2001 to 2013, for each variable. However, there are a few missing values, especially regarding the perceived opportunities for some countries, namely Estonia, Luxembourg and Slovak Republic.

**World Bank Data**

As mentioned above, one of our control variables, annual GDP growth, was obtained using the [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators) API from the World Bank.  Using this API, we accessed GDP data from all the OECD countries directly from R and R will pull this information from WDI whenever this code is run. To clean this data, we used plyr to change the variable names then recoded the U.S. and South Korea country names to maintain consistency with the GEM data set and dropped an unnecessary column that was not included with the GEM data.

Finally, we merged the annual GDP growth data with the *FinalGEM* data set to create *CompleteDataset*.  This is the data set used to run the regressions and generate the charts and figures outlined in this report.  To reduce complexity in the environment we also included a command to remove unnecessary objects from the environment (all the individual original and cleaned GEM data, and the intermediate merged data sets used to create *FinalGEM*)

<br>
```{r echo =FALSE, cache=TRUE, include=FALSE, results='asis'}

@@

# Library required packages repmis, ridy, reshape2, plyr, WDI, and stargazer packages
library(repmis)
library(tidyr)
library(reshape2)
library(plyr)
library(WDI)
library(stargazer)

@@

#### Data Analysis

To review, the three regression equations run for this assignment are the following:

1. $PercentIntention = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

2. $PercentNascent = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

3. $PercentOwner = PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth$

To examine the relationship between these variables, we used least square estimation, the results of which can be found in the following table.  The package *stargazer* was required to generate this table.  As evidenced by the equation results, the individual variables selected are statistically significant, but the model only explains a relatively small degree of the variation in the dependent variable.  For example, the adjusted $R^2$ of regression equation one is .29, the adjusted $R^2$ of regression equation two is .38, and the adjusted $R^2$ is .13.  
<br>

# Run linear regression R1
R1 <- lm(PercentIntention ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

@@

# Run linear regression R2
R2 <- lm(PercentNascent ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

@@

# Run linear regression R3
R3 <- lm(PercentOwner ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteDataset)

# Obtain summary statistics of R3
summary(R3)
```

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis'}

# Create regression output table
labels <- c('Percent Perceived', 'Percent Fear',
            'Annual GDP Growth', 'Percent Opportunities', 'Percent Fear x Percent Opportunities','(Intercept)')

stargazer::stargazer(R1, R2, R3, covariate.labels = labels, 
                     title = 'Linear Regression Estimates', digits = 2, type = 'html')
```
<br>

**Simulation Package and Cleaning**

In addition to the output table, a simulation is helpful to visually illustrate the relationship the probability of obtaining a given value in the dependent variables conditional on change in the independent variable.  We used the *Zelig* package to run a least square regression and generate a simulation of equation 1.  Prior to running this regression, however, we ran the na.omit command to remove omitted variables from the equation.

To examine the probabilities of achieving given outcomes, we generated three simulations for the first regression equation.  To demonstrate this method, we have included one simulation below.

**Process**

First, we estimated a least square equation with variation in the variable *PercentPerceived* of 20-100 percent. The ranges for variance in the independent variables was chosen by identifying the best range for capturing the largest degree of variance for the model.  The first simulation is below.  The y axis shows variation in *PercentIntention* conditional on change in *PercentPerceived* and the x axis shows variation in *PercentPerceived*. The results may be interpreted as showing that people who believe they have the knowledge required to start a business may or may not do so (with a high degree of variation), while people who do not believe they have this knowledge are consistently unlikely to start a business (there is little variation).

```{r echo = FALSE, cache=TRUE, include=FALSE, results='asis'}

# Library Zelig (can be buggy if loaded to early)
library(Zelig)

# Omit omitted data from simulation
CompleteData_omit <- na.omit(CompleteDataset)

# Run a least squares regression with Zelig
Z1 <- zelig(PercentIntention ~ PercentPerceived + PercentFear*PercentOpportunities + AnnualGDPGrowth, data = CompleteData_omit, model = "ls")
```

```{r echo = FALSE, cache=TRUE, include=TRUE, results='asis' }

# Set range of PercentPerceived values in simulation
x.out2 <- setx(Z1, PercentPerceived = 20:100)

# Run simulation (default 1000 trials)
simz2 <- sim(Z1, x = x.out2)

# Adjust margins
par(mar=c(2,2,2,2))

# Plot simulation
plot(simz2)

@@

<br>
